### 2022-06-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a8f5b9e565b086cbaef1eee76be468c67b3a2d0e 🚀

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{28,47} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species\textsuperscript{26,48,49} (add cite for
10.1038/nature13139). This represents a challenge for both conservation
strategies and pandemic prevention,\textsuperscript{50} but identifying
areas at risk, and protecting the health of bats and ecosystems within
those zones, can be a win-win intervention for both (add cites for DOIs:
10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{51} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{52} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{53} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{54} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{55} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{56} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{57} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{58} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{59} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{60} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{61} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{62} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
likelihood tree reconstructed in IQ-TREE\textsuperscript{63} v1.6.12
with ModelFinder\textsuperscript{64} ultrafast bootstrap
approximation\textsuperscript{65} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{66} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{67,68} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{69} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{51. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}

\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
Abstraction software Library}. (Zenodo, 2022).
doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}

\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}

\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
mammal tree: Species-level sets of phylogenies for questions in ecology,
evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
(2019).}

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
variance of community data: Dissimilarity coefficients and partitioning.
\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}

\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
of beta diversity in the Barro Colorado Island forest dynamics plot,
Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}

\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
ecological uniqueness over broad spatial extents using species
distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}

\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
Predicting the global mammalian viral sharing network using
phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}

\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
geography of lake carbon cycling. \emph{Limnology and Oceanography
Letters} \textbf{3}, 49--56 (2018).}

\leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
\CSLLeftMargin{60. }
\CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
S. L. \& Hochberg, M. E. Hot Spots, Cold Spots, and the Geographic
Mosaic Theory of Coevolution. \emph{The American Naturalist}
\textbf{156}, 156--174 (2000).}

\leavevmode\hypertarget{ref-Keke2010StuSki}{}%
\CSLLeftMargin{61. }
\CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
image segmentation used by Fuzzy-c-means arithmetic. in \emph{2010
Seventh International Conference on Fuzzy Systems and Knowledge
Discovery} vol. 2 612--615 (2010).}

\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
\CSLLeftMargin{62. }
\CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
Alignment Software Version 7: Improvements in Performance and Usability.
\emph{Molecular Biology and Evolution} \textbf{30}, 772--780 (2013).}

\leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
\CSLLeftMargin{63. }
\CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
B. Q. IQ-TREE: A Fast and Effective Stochastic Algorithm for Estimating
Maximum-Likelihood Phylogenies. \emph{Molecular Biology and Evolution}
\textbf{32}, 268--274 (2015).}

\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
\CSLLeftMargin{64. }
\CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
Haeseler, A. \& Jermiin, L. S. ModelFinder: Fast model selection for
accurate phylogenetic estimates. \emph{Nature Methods} \textbf{14},
587--589 (2017).}

\leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
\CSLLeftMargin{65. }
\CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
Q. \& Vinh, L. S. UFBoot2: Improving the Ultrafast Bootstrap
Approximation. \emph{Molecular Biology and Evolution} \textbf{35},
518--522 (2018).}

\leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
\CSLLeftMargin{66. }
\CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
\& Baillie, J. E. M. Mammals on the EDGE: Conservation Priorities Based
on Threat and Phylogeny. \emph{PLOS ONE} \textbf{2}, e296 (2007).}

\leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
\CSLLeftMargin{67. }
\CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
of vascular plant diversity. \emph{Proceedings of the National Academy
of Sciences} \textbf{104}, 5925--5930 (2007).}

\leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
\CSLLeftMargin{68. }
\CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
biogeographical regions based on species distributions. \emph{Journal of
Biogeography} \textbf{37}, 2029--2053 (2010).}

\leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
\CSLLeftMargin{69. }
\CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for
modern phylogenetics and evolutionary analyses in R.
\emph{Bioinformatics} \textbf{35}, 526--528 (2019).}

>>>>

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{28,47} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{26,48--50} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{51} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both.\textsuperscript{52,53}

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{54} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{56} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{57} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{59} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{60} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{61} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{62} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{63} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{64} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{65} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
likelihood tree reconstructed in IQ-TREE\textsuperscript{66} v1.6.12
with ModelFinder\textsuperscript{67} ultrafast bootstrap
approximation\textsuperscript{68} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{69} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{70,71} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{72} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
E. A. \& Daily, G. C. Predicting biodiversity change and averting
collapse in agricultural landscapes. \emph{Nature} \textbf{509},
213--217 (2014).}

\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
\CSLLeftMargin{51. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
interventions that benefit human health and conservation. \emph{Nature
Sustainability} \textbf{4}, 298--304 (2021).}

\leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
spillover: A call to action to safeguard environmental, animal, and
human health. \emph{The Lancet Planetary Health} \textbf{5}, e237--e245
(2021).}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{54. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}

\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
Abstraction software Library}. (Zenodo, 2022).
doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}

\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}

\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
mammal tree: Species-level sets of phylogenies for questions in ecology,
evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
(2019).}

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
variance of community data: Dissimilarity coefficients and partitioning.
\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}

\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
of beta diversity in the Barro Colorado Island forest dynamics plot,
Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}

\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
\CSLLeftMargin{60. }
\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
ecological uniqueness over broad spatial extents using species
distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}

\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
\CSLLeftMargin{61. }
\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
Predicting the global mammalian viral sharing network using
phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}

\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
\CSLLeftMargin{62. }
\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
geography of lake carbon cycling. \emph{Limnology and Oceanography
Letters} \textbf{3}, 49--56 (2018).}

\leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
\CSLLeftMargin{63. }
\CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
S. L. \& Hochberg, M. E. Hot Spots, Cold Spots, and the Geographic
Mosaic Theory of Coevolution. \emph{The American Naturalist}
\textbf{156}, 156--174 (2000).}

\leavevmode\hypertarget{ref-Keke2010StuSki}{}%
\CSLLeftMargin{64. }
\CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
image segmentation used by Fuzzy-c-means arithmetic. in \emph{2010
Seventh International Conference on Fuzzy Systems and Knowledge
Discovery} vol. 2 612--615 (2010).}

\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
\CSLLeftMargin{65. }
\CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
Alignment Software Version 7: Improvements in Performance and Usability.
\emph{Molecular Biology and Evolution} \textbf{30}, 772--780 (2013).}

\leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
\CSLLeftMargin{66. }
\CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
B. Q. IQ-TREE: A Fast and Effective Stochastic Algorithm for Estimating
Maximum-Likelihood Phylogenies. \emph{Molecular Biology and Evolution}
\textbf{32}, 268--274 (2015).}

\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
\CSLLeftMargin{67. }
\CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
Haeseler, A. \& Jermiin, L. S. ModelFinder: Fast model selection for
accurate phylogenetic estimates. \emph{Nature Methods} \textbf{14},
587--589 (2017).}

\leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
\CSLLeftMargin{68. }
\CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
Q. \& Vinh, L. S. UFBoot2: Improving the Ultrafast Bootstrap
Approximation. \emph{Molecular Biology and Evolution} \textbf{35},
518--522 (2018).}

\leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
\CSLLeftMargin{69. }
\CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
\& Baillie, J. E. M. Mammals on the EDGE: Conservation Priorities Based
on Threat and Phylogeny. \emph{PLOS ONE} \textbf{2}, e296 (2007).}

\leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
\CSLLeftMargin{70. }
\CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
of vascular plant diversity. \emph{Proceedings of the National Academy
of Sciences} \textbf{104}, 5925--5930 (2007).}

\leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
\CSLLeftMargin{71. }
\CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
biogeographical regions based on species distributions. \emph{Journal of
Biogeography} \textbf{37}, 2029--2053 (2010).}

\leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
\CSLLeftMargin{72. }
\CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for
modern phylogenetics and evolutionary analyses in R.
\emph{Bioinformatics} \textbf{35}, 526--528 (2019).}
